Puck Predictions: Unraveling the NHL Game Forecasting Riddle

### 2024-04-10 Latex/draft.tex
Change tables from 'figure' to 'table.

To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
then the away team shot the puck and missed, the markov chain would look like Figure~\ref{fig:markov_chain_sample}.

\begin{figure}
    \centering
    \begin{tabular}{ |c|c| } 
    \hline
        \multicolumn{2}{|c|}{\textbf{Sample Markov Chain}} \\
     \hline
     \textit{Next Event} & \textit{Transition Probability} \\ 
     \hline
     Shot Away & 0.1853207 \\ 
     \hline
     Blocked Shot Away & 0.1138666 \\ 
     \hline
     \vdots & \vdots \\
    \hline
    Penalty Away & 0.0167210 \\
    \hline
    Goal Home & 0.0085793 \\
    \hline
    \end{tabular}
    \caption{The transition events and probabilities when the previous three events were Faceoff Home, Giveaway Home, and Missed
    Shot Away, respectively.}
    \label{fig:markov_chain_sample}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tabular}{ |c|c|c|c| } 
    \hline
        \multicolumn{4}{|c|}{\textbf{Accuracy Comparison Between Models}} \\
     \hline
     \textit{Game Seconds Played} & \textit{MCMC} & \textit{XGBoost} & \textit{Bayesian} \\ 
     \hline
     0 & a & a & a \\ 
     \hline
     600 & a & a & a \\ 
     \hline
     1200 & a & a & a \\
    \hline
    1800 & a & a & a \\
    \hline
    2400 & a & a & a \\
    \hline
    3000 & a & a & a \\
    \hline
    3600 & a & a & a \\
    \hline
    \end{tabular}
    \caption{Each of our model's predictive accuracy over the course of 100 games}
    \label{fig:game_accuracies}
\end{figure}

@@

In Figure~\ref{fig:game_accuracies} we note that each model increased in accuracy as the game progressed. We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate. We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set. Also, during experimentation with the models, we noticed that the MCMC simulator and the Bayesian Network tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning. The MCMC and Bayesian Network are structured to model the flow of events, so it appears that they picked up on a more predictive flow in how away teams play.

>>>>

To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
then the away team shot the puck and missed, the markov chain would look like Table~\ref{tab:markov_chain_sample}.

\begin{table}
    \centering
    \begin{tabular}{ |c|c| } 
    \hline
        \multicolumn{2}{|c|}{\textbf{Sample Markov Chain}} \\
     \hline
     \textit{Next Event} & \textit{Transition Probability} \\ 
     \hline
     Shot Away & 0.1853207 \\ 
     \hline
     Blocked Shot Away & 0.1138666 \\ 
     \hline
     \vdots & \vdots \\
    \hline
    Penalty Away & 0.0167210 \\
    \hline
    Goal Home & 0.0085793 \\
    \hline
    \end{tabular}
    \caption{The transition events and probabilities when the previous three events were Faceoff Home, Giveaway Home, and Missed
    Shot Away, respectively.}
    \label{tab:markov_chain_sample}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabular}{ |c|c|c|c| } 
    \hline
        \multicolumn{4}{|c|}{\textbf{Accuracy Comparison Between Models}} \\
     \hline
     \textit{Game Seconds Played} & \textit{MCMC} & \textit{XGBoost} & \textit{Bayesian} \\ 
     \hline
     0 & a & a & a \\ 
     \hline
     600 & a & a & a \\ 
     \hline
     1200 & a & a & a \\
    \hline
    1800 & a & a & a \\
    \hline
    2400 & a & a & a \\
    \hline
    3000 & a & a & a \\
    \hline
    3600 & a & a & a \\
    \hline
    \end{tabular}
    \caption{Each of our model's predictive accuracy over the course of 100 games}
    \label{tab:game_accuracies}
\end{table}

@@

In Figure~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed. We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate. We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set. Also, during experimentation with the models, we noticed that the MCMC simulator and the Bayesian Network tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning. The MCMC and Bayesian Network are structured to model the flow of events, so it appears that they picked up on a more predictive flow in how away teams play.
